---
title: "Chitra_Project_Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r import}
library(data.table)
library(dplyr)
library(ggplot2)
library(stargazer)
library(lmtest)
library(sandwich)
library(foreign)
library(knitr)
```

## Load Data

You can also embed plots, for example:

```{r loadData}
dt <- fread("data_as_of_july_25.csv")

setnames(dt, "Timestamp", "timestamp")
setnames(dt, "Name", "name")
setnames(dt, "Gender", "gender")
setnames(dt, "Age", "age")
setnames(dt, "Music Training", "music_training")
setnames(dt, "Pre-treatment Score", "pre_treatment_score")
setnames(dt, "Post-treatment Score", "post_treatment_score")
setnames(dt, "Treatment", "treatment")
setnames(dt, "Email", "email")

dt[3 : 9]
```

## Data Cleanup

1. Remove duplicate data

``` {r}
cat('Total number of rows: ', dt[,.N], '\n')
duplicate_data <- sum(duplicated(dt, by=c('name', 'gender', 'age', 'music_training','pre_treatment_score', 'Time of Day', 'post_treatment_score', 'treatment', 'email' )))
cat('Number of duplicated rows: ', duplicate_data, '\n')

d <- unique(dt, by = c('name', 'gender', 'age', 'music_training','pre_treatment_score', 'Time of Day', 'post_treatment_score', 'treatment', 'email' ))
cat('Number of unique rows in dataset: ', d[,.N], '\n')
d <- d[, 3:9]

# Randomization Check - percentage of dataset in treatment
d[age == '5 - 21 years' & treatment == 1, .N]/d[age == '5 - 21 years', .N]
d[age == '22 - 55 years' & treatment == 1, .N]/d[age == '22 - 55 years', .N]
d[age == 'Older than 55 years' & treatment == 1, .N]/d[age == 'Older than 55 years' , .N]
```

2. Remove any rows that are incomplete with data. E.g some rows dont have pre treatment or post treatment score because user made an error in data entry

``` {r}
# Eliminate all rows where pre_treatment score is 0 as these rows are not usable for data analysis
d <- new_d[pre_treatment_score != 0, ]
d[, .N]
```

3. Add additional derived columns to help analysis

```{r}
d[ , score_difference := post_treatment_score - pre_treatment_score]
d[, score_improved := ifelse (score_difference >= 0, 1, 0)]
```

```{r HelperMethods}
reg <- function (model, title = '', ptest = TRUE) {
  test <- coeftest(model, vcovHC(model))
  if(ptest) {
    print(test)
    stargazer(test[,], title = title, type='text')
  }
  return (list('model' = model, 'test' = test))
}
```

## Randomization Check

```{r}
# Check the number for whose score_diff is >= 0
cat('People in treatment for whom score increased or stayed same: ', d[treatment == 1 & score_difference >= 0, .N], '\n')
cat('People in treatment for whom score decreased: ', d[treatment == 1 & score_difference < 0, .N] , '\n')
cat('People in control for whom score increased increased or stayed same: ', d[treatment == 0 & score_difference >= 0, .N] , '\n')
cat('People in control for whom score decreased: ', d[treatment == 0 & score_difference < 0, .N], '\n')
```

## Initial Analysis

### Placebo test 

Check if the treatment has an effect on the pre treatment score. Treatment should not have any impact on it.
```{r}
m0 <- reg(d[,lm(pre_treatment_score ~ treatment)], title = 'Placebo test Estimates of pre treatment score')
```

Treatment has positive effect on pre treatment score but the p-value is not statistically significant which indicates that placebo test passes
 
## Using Score Improvement as Outcome Variable

### Basic Model

```{r scoreImprovement, warning = FALSE}
m1 <- reg(d[, lm(score_improved ~ treatment)], title = 'Basic Model:  score improvement ~ treatment')
stargazer(m1$model, type='text')
```

### Using baseline scores as covariate

```{r scoreImprovement, warning = FALSE}
m1 <- reg(d[, lm(score_improved ~ treatment + pre_treatment_score)], title = 'Basic Model: Score improvement ~ treatment + pre_treatment_score')
stargazer(m1$model, type='text')
```

## Using post_treatment_score as Outcome Variable

```{r warning=FALSE}
# using only treatment

m11 <- reg(d[, lm(post_treatment_score ~ treatment)], title = 'Basic Model with Post Treatment Score')
ate = d[treatment == 1, mean(post_treatment_score)] - d[treatment == 0, mean(post_treatment_score)]
cat('ATE: ', ate, '\n')

hist(
  d[,pre_treatment_score],
  col = 'black',
  xlab = 'Pre treatment Score',
  xlim = c(0, 12),
  main = 'Distribution of pre-treatment score',
)
abline(v = d[treatment == 1, mean(pre_treatment_score)], col = 'blue', lwd = 3)
abline(v = d[treatment == 0, mean(pre_treatment_score)], col = 'red', lwd = 3)

hist(
  d[,post_treatment_score],
  col = 'black',
  xlab = 'Post Treatment Score',
  xlim = c(0, 12),
  main = 'Distribution of post-treatment score',
)
abline(v = d[treatment == 1, mean(post_treatment_score)], col = 'blue', lwd = 3)
abline(v = d[treatment == 0, mean(post_treatment_score)], col = 'red', lwd = 3)

hist(
  d[,score_difference],
  col = 'black',
  xlab = 'Score Difference',
  xlim = c(0, 12),
  main = 'Distribution of score difference',
)
abline(v = d[treatment == 1, mean(score_difference)], col = 'blue', lwd = 3)
abline(v = d[treatment == 0, mean(score_difference)], col = 'red', lwd = 3)

#m1 <- reg(d[age == '5 - 21 years',lm(post_treatment_score ~ treatment)], ptest = FALSE)
#m2 <- reg(d[age == '22 - 55 years',lm(post_treatment_score ~ treatment)], ptest = FALSE)
#m3 <- reg(d[age == 'Older than 55 years',lm(post_treatment_score ~ treatment)], ptest = FALSE)
# stargazer( m1$model, m2$model, m3$model,
#  type = 'text',
#  title = 'Simple Treatment Effect of receiving a treatment',
#  se = list(sqrt(diag(vcovHC(m1$model))), sqrt(diag(vcovHC(m2$model))), sqrt(diag(vcovHC(m3$model)))),
#  add.lines = list(c("Age Group:", "5 - 21 years", "22 - 55 years", "Older than 55 years")))

m1 <- reg(d[, lm(post_treatment_score ~ treatment)], title = 'Basic Model with Post Treatment Score')
stargazer( m1$model, 
 type = 'text',
 title = 'Simple Treatment Effect of receiving a treatment',
 se = list(sqrt(diag(vcovHC(m1$model)))))

```

### Basic Model 2 - Pretreament score + treatment

```{r warning=FALSE}
m0 <- reg(d[,lm(post_treatment_score ~  treatment + pre_treatment_score)], ptest = FALSE)
stargazer( m0$model,
  type = 'text',
  title = 'Treatment Effect of receiving a treatment with pre treatment score',
  se = list(sqrt(diag(vcovHC(m0$model))))
)

# m1 <- reg(d[age == '5 - 21 years',lm(post_treatment_score ~ treatment + pre_treatment_score)], ptest = FALSE)
# m2 <- reg(d[age == '22 - 55 years',lm(post_treatment_score ~ treatment + pre_treatment_score)], ptest = FALSE)
# m3 <- reg(d[age == 'Older than 55 years',lm(post_treatment_score ~ treatment + pre_treatment_score)], ptest = FALSE)
# stargazer( m1$model, m2$model, m3$model,
#   type = 'text',
#   title = 'Treatment Effect of receiving a treatment with pre treatment score',
#   se = list(sqrt(diag(vcovHC(m1$model))), sqrt(diag(vcovHC(m2$model))), sqrt(diag(vcovHC(m3$model)))),
#   add.lines = list(c("Age Group:", "5 - 21 years", "22 - 55 years", "Older than 55 years")))
```

### Block average as covariate

```{r warning=FALSE}
d[, avg_block_turnout := mean(post_treatment_score), by=age]
head(d)

# below gives highly significant results for the pre-treatment scores
m0 <- reg(d[,lm(post_treatment_score ~  treatment + avg_block_turnout)], ptest = TRUE) 
stargazer( m0$model,
  type = 'text',
  title = 'Treatment Effect of receiving a treatment with pre treatment score',
  se = list(sqrt(diag(vcovHC(m0$model))))
)

# m0 <- reg(d[,lm(score_difference ~  treatment + pre_treatment_score + avg_block_turnout)], ptest = TRUE)
# m1 <- reg(d[age == '5 - 21 years',lm(score_difference ~ treatment + pre_treatment_score + avg_block_turnout)], ptest = FALSE)
# m2 <- reg(d[age == '22 - 55 years',lm(score_difference ~ treatment + pre_treatment_score + avg_block_turnout)], ptest = FALSE)
# m3 <- reg(d[age == 'Older than 55 years',lm(score_difference ~ treatment + pre_treatment_score  + avg_block_turnout)], ptest = FALSE)
# stargazer( m1$model, m2$model, m3$model,
#   type = 'text',
#   title = 'Treatment Effect of receiving a treatment with block fixed effects',
#   se = list(sqrt(diag(vcovHC(m1$model))), sqrt(diag(vcovHC(m2$model))), sqrt(diag(vcovHC(m3$model)))),
#   add.lines = list(c("Age Group:", "5 - 21 years", "22 - 55 years", "Older than 55 years")))

```
```{r warning=FALSE}
unique(d[,music_training])
d[,music := ifelse(music_training == 'more than 5 years of training', 1, 0)]
m3 <- reg(d[,lm(post_treatment_score ~  treatment + avg_block_turnout + music + treatment * music )])
stargazer( m3$model,
  type = 'text',
  title = 'Treatment Effect of receiving a treatment with pre treatment score',
  se = list(sqrt(diag(vcovHC(m3$model))))
)
```

