---
title: "Chitra_Project_Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r import}
library(data.table)
library(dplyr)
library(ggplot2)
library(stargazer)
library(lmtest)
library(sandwich)
library(foreign)
library(knitr)
```

## Load Data

You can also embed plots, for example:

```{r loadData}
#dt <- fread("data_as_of_july_25.csv")
dt <- fread("data_26.csv")

setnames(dt, "Timestamp", "timestamp")
setnames(dt, "Name", "name")
setnames(dt, "Gender", "gender")
setnames(dt, "Age", "age")
setnames(dt, "Music Training", "music_training")
setnames(dt, "Pre-treatment Score", "pre_treatment_score")
setnames(dt, "Post-treatment Score", "post_treatment_score")
setnames(dt, "Treatment", "treatment")
setnames(dt, "Email", "email")
setnames(dt, 'Time of Day', 'time_of_day')

dt[, 3 : 9]
dt[age == '5 - 21 years',]
```

## Data Cleanup

1. Remove duplicate data

``` {r}
cat('Total number of rows: ', dt[,.N], '\n')
duplicate_data <- sum(duplicated(dt, by=c('name', 'gender', 'age', 'music_training','pre_treatment_score', 'time_of_day', 'post_treatment_score', 'treatment', 'email' )))
cat('Number of duplicated rows: ', duplicate_data, '\n')

d <- unique(dt, by = c('name', 'gender', 'age', 'music_training','pre_treatment_score', 'time_of_day', 'post_treatment_score', 'treatment', 'email' ))
cat('Number of unique rows in dataset: ', d[,.N], '\n')
d <- d[, 3:9]

# Randomization Check - percentage of dataset in treatment
d[age == '5 - 21 years' & treatment == 1, .N]/d[age == '5 - 21 years', .N]
d[age == '22 - 55 years' & treatment == 1, .N]/d[age == '22 - 55 years', .N]
d[age == 'Older than 55 years' & treatment == 1, .N]/d[age == 'Older than 55 years' , .N]
```

2. Remove any rows that are incomplete with data. E.g some rows dont have pre treatment or post treatment score because user made an error in data entry

``` {r}
# Eliminate all rows where pre_treatment score is 0 as these rows are not usable for data analysis
d[pre_treatment_score == 0, ]
d <- d[pre_treatment_score > 0, ]
##d[post_treatment_score == 2, .N]
d <- d[post_treatment_score > 2, ]
d[, .N]
```

3. Add additional derived columns to help analysis

```{r}
d[ , score_difference := post_treatment_score - pre_treatment_score]
d[, score_improved := ifelse (score_difference >= 0, 1, 0)]

# Convert columns to factors where appropriate
d[, treatment := as.factor(treatment)]
d[, age := as.factor(age)]
d[, music_training := as.factor(music_training)]
d[, gender := as.factor(gender)]
d[, time_of_day := as.factor(time_of_day)]
d
```

```{r HelperMethods}
reg <- function (model, title = '', ptest = TRUE) {
  test <- coeftest(model, vcovHC(model))
  if(ptest) {
    print(test)
    stargazer(test[,], title = title, type='text')
  }
  return (list('model' = model, 'test' = test))
}
```

## EDA

```{r}
library(gridExtra)
# Draw box plots of treatment and control for each age block
plot1 <- ggplot(d, aes(x=treatment, y=post_treatment_score)) + geom_boxplot() + ggtitle('All subjects')

# Draw by Age
plot2 <- ggplot(d[age == '5 - 21 years',], aes(x=treatment, y=post_treatment_score)) + geom_boxplot() + ggtitle('5 - 21 Years')
plot3 <- ggplot(d[age == '22 - 55 years',], aes(x=treatment, y=post_treatment_score)) + geom_boxplot() + ggtitle('22 - 55 years')
plot4 <- ggplot(d[age == 'Older than 55 years',], aes(x=treatment, y=post_treatment_score)) + geom_boxplot() + ggtitle('Older than 55 years')
grid.arrange(plot1 , plot2, plot3 , plot4, ncol = 2, top = 'Box Plot of Post Treatment Score')
```

```{r}
library(gridExtra)
# Draw box plots of treatment and control for each age block
plot1 <- ggplot(d, aes(x=treatment, y=score_difference)) + geom_boxplot() + ggtitle('All subjects')

# Draw by Age
plot2 <- ggplot(d[age == '5 - 21 years',], aes(x=treatment, y=score_difference)) + geom_boxplot() + ggtitle('5 - 21 Years')
plot3 <- ggplot(d[age == '22 - 55 years',], aes(x=treatment, y=score_difference)) + geom_boxplot() + ggtitle('22 - 55 years')
plot4 <- ggplot(d[age == 'Older than 55 years',], aes(x=treatment, y=score_difference)) + geom_boxplot() + ggtitle('Older than 55 years')
grid.arrange(plot1 , plot2, plot3 , plot4, ncol = 2, top = 'Box Plot of Score Difference')
```

## Randomization Check

```{r}
# Check the number for whose score_diff is >= 0
cat('People in treatment for whom score increased or stayed same: ', d[treatment == 1 & score_difference >= 0, .N], '\n')
cat('People in treatment for whom score decreased: ', d[treatment == 1 & score_difference < 0, .N] , '\n')
cat('People in control for whom score increased or stayed same: ', d[treatment == 0 & score_difference >= 0, .N] , '\n')
cat('People in control for whom score decreased: ', d[treatment == 0 & score_difference < 0, .N], '\n')
```

## Initial Analysis

### Power Calculation

``` {r}
N = 125

mu_t <- mean(d[treatment == 1, score_improved])
mu_c <- mean(d[treatment == 0, score_improved])
mu <- mu_t - mu_c
sigma <- sqrt(var(d[, score_improved]))

power <- pnorm((abs(mu)*sqrt(N))/(2*sigma) - qnorm(0.975))
power

mu_t <- mean(d[treatment == 1, score_difference])
mu_c <- mean(d[treatment == 0, score_difference])
mu <- mu_t - mu_c
sigma <- sqrt(var(d[, score_difference]))
power <- pnorm((abs(mu)*sqrt(N))/(2*sigma) - qnorm(0.975))
power

mu_t <- mean(d[treatment == 1, post_treatment_score])
mu_c <- mean(d[treatment == 0, post_treatment_score])
mu <- mu_t - mu_c
sigma <- sqrt(var(d[, post_treatment_score]))
power <- pnorm((abs(mu)*sqrt(N))/(2*sigma) - qnorm(0.975))
power
```

### Placebo test 

Check if the treatment has an effect on the pre treatment score. Treatment should not have any impact on it.
```{r}
m0 <- reg(d[,lm(pre_treatment_score ~ treatment)], title = 'Placebo test Estimates of pre treatment score')
```

Treatment has positive effect on pre treatment score but the p-value is not statistically significant which indicates that placebo test passes
 
## Using Score Improvement as Outcome Variable


### Basic Model

```{r scoreImprovement1, warning = FALSE}
m0 <- reg(d[, lm(score_improved ~ treatment)], title = 'Basic Model:  score improvement ~ treatment')
#stargazer(m1$model, type='text')


m1 <- reg(d[age == '5 - 21 years',lm(score_improved ~ treatment)], ptest = FALSE)
m2 <- reg(d[age == '22 - 55 years',lm(score_improved ~ treatment)], ptest = FALSE)
m3 <- reg(d[age == 'Older than 55 years',lm(score_improved ~ treatment)], ptest = FALSE)
stargazer( m0$model, m1$model, m2$model, m3$model,
 type = 'text',
 title = 'Simple Treatment Effect of receiving a treatment',
 se = list(sqrt(diag(vcovHC(m0$model))), sqrt(diag(vcovHC(m1$model))), sqrt(diag(vcovHC(m2$model))), sqrt(diag(vcovHC(m3$model)))),
 column.labels = c("All ages", "5 - 21 years", "22 - 55 years", "Older than 55 years"),
 add.lines = list(c("Age Group:", "All", "5 - 21 years", "22 - 55 years", "Older than 55 years")),
 dep.var.caption  = "Outcome Variable: score-improved",
 dep.var.labels.include = FALSE
 )
```

### Using baseline scores as covariate

```{r scoreImprovement2, warning = FALSE}
m0 <- reg(d[, lm(score_improved ~ treatment + pre_treatment_score)], title = 'Basic Model: Score improvement ~ treatment + pre_treatment_score')
#stargazer(m1$model, type='text')

m1 <- reg(d[age == '5 - 21 years',lm(score_improved ~ treatment + pre_treatment_score)], ptest = FALSE)
m2 <- reg(d[age == '22 - 55 years' ,lm(score_improved ~ treatment + pre_treatment_score)], ptest = FALSE)
m3 <- reg(d[age == 'Older than 55 years' ,lm(score_improved ~ treatment + pre_treatment_score)], ptest = FALSE)
stargazer( m0$model, m1$model, m2$model, m3$model,
 type = 'text',
 title = 'Simple Treatment Effect of receiving a treatment2',
 se = list(sqrt(diag(vcovHC(m0$model))),sqrt(diag(vcovHC(m1$model))), sqrt(diag(vcovHC(m2$model))), sqrt(diag(vcovHC(m3$model)))),
 column.labels = c("All ages", "5 - 21 years", "22 - 55 years", "Older than 55 years"),
 add.lines = list(c("Model: ", "All ages", "5 - 21 years", "22 - 55 years", "Older than 55 years")),
 dep.var.caption  = "Outcome Variable: score-improved",
 dep.var.labels.include = FALSE
)
```

### Adding music

```{r warning=FALSE}
unique(d[,music_training])
d[,music := ifelse(music_training == 'more than 5 years of training', 1, 0)]
#d[,music := ifelse(music_training == 'No training' , 0, 1)]
m3 <- reg(d[,lm(score_improved ~  treatment + pre_treatment_score + music + treatment * music)])
stargazer( m3$model,
  type = 'text',
  title = 'Treatment Effect of receiving a treatment with pre treatment score',
  se = list(sqrt(diag(vcovHC(m3$model))))
)
```

### Time of Day

```{r warning=FALSE}
d[ , time := time_of_day]
d[time_of_day == '4 AM - 8 AM' | time_of_day == 'Midnight - 4 AM', time := 'Midnight to 8AM']
d[, .(mean_score_diff = mean(score_difference), mean_score_improved = mean(score_improved)), by = time]
```

```{r warning=FALSE}
#d[, morning := ifelse(time == '8 AM - Noon' | time == 'Noon - 4 PM', 1, 0)]
m3 <- reg(d[,lm(score_improved ~  treatment + pre_treatment_score + time + treatment * time)])
stargazer( m3$model,
  type = 'text',
  title = 'Treatment Effect of receiving a treatment at specific time of the day',
  se = list(sqrt(diag(vcovHC(m3$model))))
)
```

## Full Model

```{r warning=FALSE}

m3 <- reg(d[,lm(score_improved ~  treatment + pre_treatment_score + music + time )])
stargazer( m3$model,
  type = 'text',
  title = 'Treatment Effect of receiving a treatment at specific time of the day',
  se = list(sqrt(diag(vcovHC(m3$model))))
)
```

## Using score_differnece as Outcome Variable

```{r warning=FALSE}
# using only treatment

m11 <- reg(d[, lm(score_difference ~ treatment)], title = 'Basic Model with Score Difference as outcome variable')
ate = d[treatment == 1, mean(score_difference)] - d[treatment == 0, mean(score_difference)]
cat('ATE: ', ate, '\n')

hist(
  d[,score_difference],
  col = 'black',
  xlab = 'Score Difference',
  xlim = c(0, 12),
  main = 'Distribution of score difference',
)
abline(v = d[treatment == 1, mean(score_difference)], col = 'blue', lwd = 3)
abline(v = d[treatment == 0, mean(score_difference)], col = 'red', lwd = 3)

#m1 <- reg(d[age == '5 - 21 years',lm(post_treatment_score ~ treatment)], ptest = FALSE)
#m2 <- reg(d[age == '22 - 55 years',lm(post_treatment_score ~ treatment)], ptest = FALSE)
#m3 <- reg(d[age == 'Older than 55 years',lm(post_treatment_score ~ treatment)], ptest = FALSE)
# stargazer( m1$model, m2$model, m3$model,
#  type = 'text',
#  title = 'Simple Treatment Effect of receiving a treatment',
#  se = list(sqrt(diag(vcovHC(m1$model))), sqrt(diag(vcovHC(m2$model))), sqrt(diag(vcovHC(m3$model)))),
#  add.lines = list(c("Age Group:", "5 - 21 years", "22 - 55 years", "Older than 55 years")))

m1 <- reg(d[, lm(score_difference ~ treatment)], title = 'Basic Model with Score Difference')
stargazer( m1$model, 
 type = 'text',
 title = 'Simple Treatment Effect of receiving a treatment',
 se = list(sqrt(diag(vcovHC(m1$model)))))

```

### Basic Model 2 - Pretreament score + treatment

```{r warning=FALSE}
m0 <- reg(d[,lm(score_difference ~  treatment + pre_treatment_score)], ptest = FALSE)
stargazer( m0$model,
  type = 'text',
  title = 'Treatment Effect of receiving a treatment with pre treatment score',
  se = list(sqrt(diag(vcovHC(m0$model))))
)

# m1 <- reg(d[age == '5 - 21 years',lm(post_treatment_score ~ treatment + pre_treatment_score)], ptest = FALSE)
# m2 <- reg(d[age == '22 - 55 years',lm(post_treatment_score ~ treatment + pre_treatment_score)], ptest = FALSE)
# m3 <- reg(d[age == 'Older than 55 years',lm(post_treatment_score ~ treatment + pre_treatment_score)], ptest = FALSE)
# stargazer( m1$model, m2$model, m3$model,
#   type = 'text',
#   title = 'Treatment Effect of receiving a treatment with pre treatment score',
#   se = list(sqrt(diag(vcovHC(m1$model))), sqrt(diag(vcovHC(m2$model))), sqrt(diag(vcovHC(m3$model)))),
#   add.lines = list(c("Age Group:", "5 - 21 years", "22 - 55 years", "Older than 55 years")))
```

### Block average as covariate

```{r warning=FALSE}
d[, avg_block_turnout := mean(post_treatment_score), by=age]
#d[, avg_block_diff := mean(score_difference), by=age]
head(d)

# below gives highly significant results for the pre-treatment scores
m0 <- reg(d[,lm(score_difference ~  treatment + avg_block_turnout)], ptest = TRUE) 
stargazer( m0$model,
  type = 'text',
  title = 'Treatment Effect of receiving a treatment with pre treatment score',
  se = list(sqrt(diag(vcovHC(m0$model))))
)

# m0 <- reg(d[,lm(score_difference ~  treatment + pre_treatment_score + avg_block_turnout)], ptest = TRUE)
# m1 <- reg(d[age == '5 - 21 years',lm(score_difference ~ treatment + pre_treatment_score + avg_block_turnout)], ptest = FALSE)
# m2 <- reg(d[age == '22 - 55 years',lm(score_difference ~ treatment + pre_treatment_score + avg_block_turnout)], ptest = FALSE)
# m3 <- reg(d[age == 'Older than 55 years',lm(score_difference ~ treatment + pre_treatment_score  + avg_block_turnout)], ptest = FALSE)
# stargazer( m1$model, m2$model, m3$model,
#   type = 'text',
#   title = 'Treatment Effect of receiving a treatment with block fixed effects',
#   se = list(sqrt(diag(vcovHC(m1$model))), sqrt(diag(vcovHC(m2$model))), sqrt(diag(vcovHC(m3$model)))),
#   add.lines = list(c("Age Group:", "5 - 21 years", "22 - 55 years", "Older than 55 years")))

```
```{r warning=FALSE}
unique(d[,music_training])
d[,music := ifelse(music_training == 'more than 5 years of training', 1, 0)]
m3 <- reg(d[,lm(score_difference ~  treatment + avg_block_turnout + music + treatment * music )])
stargazer( m3$model,
  type = 'text',
  title = 'Treatment Effect of receiving a treatment with pre treatment score',
  se = list(sqrt(diag(vcovHC(m3$model))))
)
```


## Using post_treatment_score as Outcome Variable

```{r warning=FALSE}
# using only treatment

m11 <- reg(d[, lm(post_treatment_score ~ treatment)], title = 'Basic Model with Post Treatment Score')
ate = d[treatment == 1, mean(post_treatment_score)] - d[treatment == 0, mean(post_treatment_score)]
cat('ATE: ', ate, '\n')

hist(
  d[,pre_treatment_score],
  col = 'black',
  xlab = 'Pre treatment Score',
  xlim = c(0, 12),
  main = 'Distribution of pre-treatment score',
)
abline(v = d[treatment == 1, mean(pre_treatment_score)], col = 'blue', lwd = 3)
abline(v = d[treatment == 0, mean(pre_treatment_score)], col = 'red', lwd = 3)

hist(
  d[,post_treatment_score],
  col = 'black',
  xlab = 'Post Treatment Score',
  xlim = c(0, 12),
  main = 'Distribution of post-treatment score',
)
abline(v = d[treatment == 1, mean(post_treatment_score)], col = 'blue', lwd = 3)
abline(v = d[treatment == 0, mean(post_treatment_score)], col = 'red', lwd = 3)

hist(
  d[,score_difference],
  col = 'black',
  xlab = 'Score Difference',
  xlim = c(0, 12),
  main = 'Distribution of score difference',
)
abline(v = d[treatment == 1, mean(score_difference)], col = 'blue', lwd = 3)
abline(v = d[treatment == 0, mean(score_difference)], col = 'red', lwd = 3)

#m1 <- reg(d[age == '5 - 21 years',lm(post_treatment_score ~ treatment)], ptest = FALSE)
#m2 <- reg(d[age == '22 - 55 years',lm(post_treatment_score ~ treatment)], ptest = FALSE)
#m3 <- reg(d[age == 'Older than 55 years',lm(post_treatment_score ~ treatment)], ptest = FALSE)
# stargazer( m1$model, m2$model, m3$model,
#  type = 'text',
#  title = 'Simple Treatment Effect of receiving a treatment',
#  se = list(sqrt(diag(vcovHC(m1$model))), sqrt(diag(vcovHC(m2$model))), sqrt(diag(vcovHC(m3$model)))),
#  add.lines = list(c("Age Group:", "5 - 21 years", "22 - 55 years", "Older than 55 years")))

m1 <- reg(d[, lm(post_treatment_score ~ treatment)], title = 'Basic Model with Post Treatment Score')
stargazer( m1$model, 
 type = 'text',
 title = 'Simple Treatment Effect of receiving a treatment',
 se = list(sqrt(diag(vcovHC(m1$model)))))

```

### Basic Model 2 - Pretreament score + treatment

```{r warning=FALSE}
m0 <- reg(d[,lm(post_treatment_score ~  treatment + pre_treatment_score)], ptest = FALSE)
stargazer( m0$model,
  type = 'text',
  title = 'Treatment Effect of receiving a treatment with pre treatment score',
  se = list(sqrt(diag(vcovHC(m0$model))))
)

# m1 <- reg(d[age == '5 - 21 years',lm(post_treatment_score ~ treatment + pre_treatment_score)], ptest = FALSE)
# m2 <- reg(d[age == '22 - 55 years',lm(post_treatment_score ~ treatment + pre_treatment_score)], ptest = FALSE)
# m3 <- reg(d[age == 'Older than 55 years',lm(post_treatment_score ~ treatment + pre_treatment_score)], ptest = FALSE)
# stargazer( m1$model, m2$model, m3$model,
#   type = 'text',
#   title = 'Treatment Effect of receiving a treatment with pre treatment score',
#   se = list(sqrt(diag(vcovHC(m1$model))), sqrt(diag(vcovHC(m2$model))), sqrt(diag(vcovHC(m3$model)))),
#   add.lines = list(c("Age Group:", "5 - 21 years", "22 - 55 years", "Older than 55 years")))
```

### Block average as covariate

```{r warning=FALSE}
d[, avg_block_turnout := mean(post_treatment_score), by=age]
head(d)

# below gives highly significant results for the pre-treatment scores
m0 <- reg(d[,lm(post_treatment_score ~  treatment + avg_block_turnout)], ptest = TRUE) 
stargazer( m0$model,
  type = 'text',
  title = 'Treatment Effect of receiving a treatment with pre treatment score',
  se = list(sqrt(diag(vcovHC(m0$model))))
)
```

### Adding music

```{r warning=FALSE}
unique(d[,music_training])
d[,music := ifelse(music_training == 'more than 5 years of training', 1, 0)]
m3 <- reg(d[,lm(post_treatment_score ~  treatment + pre_treatment_score + music + treatment * music )])
stargazer( m3$model,
  type = 'text',
  title = 'Treatment Effect of receiving a treatment with pre treatment score',
  se = list(sqrt(diag(vcovHC(m3$model))))
)
```

